Eco-evolutionary dynamics of finite populations from first principles

### 2023-03-29 ind_files/03_BD_processes.tex
Rephrase after proofreading

\section{Description of the process and the Master Equation}
Given a population that can contain up to $m$ different (fixed) kinds of organisms, it can be entirely characterized by specifying the number of organisms of each type (Figure \ref{fig_nD_pop_description}A). Thus, the state of the population at a given time $t$ is an $m$-dimensional \emph{vector} of the form $\mathbf{n}(t) = [n_1(t),n_2(t),\ldots,n_m(t)]$, where $n_i(t)$ is the number of individuals of type $i$ in the population at time $t$.

@@

As before, we can describe the rate of change of $P(\mathbf{n},t)$, the probability of finding the population in a state $\mathbf{n}$ at time $t$, by measuring the inflow and outflow rates. Since the population changes in units of exactly one individual (by definition of a birth-death process; see section \ref{sec_math_background}), we know that these inflow and outflow rates must only involve populations that are a single individual away from our focal population. In other words, for a population $\mathbf{n} = [n_1,\ldots,n_{m}]$, the `inflow' is from all populations of the form $[n_1,\ldots,n_{i}-1,\dots,n_{m}]$ through a birth of a type $i$ individual, and from all populations of the form $[n_1,\ldots,n_{i}+1,\dots,n_{m}]$ through the death of a type $i$ individual. Thus, we have the inflow rate
\begin{equation}
\label{nD_rate_in}
\begin{split}
R_{\textrm{in}}(\mathbf{n},t) &= \sum\limits_{j=1}^{m}b_{j}([n_1,\ldots,n_{j}-1,\ldots,n_m])P([n_1,\ldots,n_{j}-1,\ldots,n_m],t) \\
& +\sum\limits_{j=1}^{m}d_{j}([n_1,\ldots,n_{j}+1,\ldots,n_m])P([n_1,\ldots,n_{j}+1,\ldots,n_m],t)
\end{split}
\end{equation}
Outflow is through births and deaths of individuals in the population $\mathbf{n}$ itself, and thus we have:
\begin{equation}
\label{nD_rate_out}
R_{\textrm{out}}(\mathbf{n},t) = \sum\limits_{j=1}^{m}b_{j}(\mathbf{n})P(\mathbf{n},t) + \sum\limits_{j=1}^{m}d_{j}(\mathbf{n})P(\mathbf{n},t)
\end{equation}
We will now define step operators, both for notational ease and in anticipation of the system size expansion. For each $i \in \{1,\ldots,m\}$, let us define two step operators $\mathcal{E}_{i}^{\pm}$ by their action on any function $f([n_1,\ldots,n_m],t)$ as:
\begin{equation}
\label{nD_step_operators}
\mathcal{E}_{i}^{\pm}f([n_1,\ldots,n_i,\ldots,n_m],t) = f([n_1,\ldots,n_i \pm 1, \ldots n_m],t)
\end{equation}
In other words, $\mathcal{E}_{i}^{\pm}$ just changes the population through the addition or removal of one type $i$ individual. We can now the rate of change of $P(\mathbf{n},t)$ as
\begin{equation}
\label{nD_intermediate_for_M_eqn}
\frac{\partial P}{\partial t}(\mathbf{n},t) = R_{\textrm{in}}(\mathbf{n},t) - R_{\textrm{out}}(\mathbf{n},t)
\end{equation}
Substituting \eqref{nD_rate_in}, \eqref{nD_rate_out}, and \eqref{nD_step_operators} into equation \eqref{nD_intermediate_for_M_eqn}, we obtain:
\begin{equation}
\label{nD_M_eqn}
\frac{\partial P}{\partial t}(\mathbf{n},t) = \sum\limits_{j=1}^{m}\left[(\mathcal{E}_j^{-}-1)b_j(\mathbf{n})P(\mathbf{n},t) + (\mathcal{E}_j^{+}-1)d_j(\mathbf{n})P(\mathbf{n},t)\right]
\end{equation}
This is the master equation of our $m$-dimensional process.

\section{The system-size expansion}
As explained in Section \ref{sec_intuition_sys_size}, we will now assume (on ecological grounds) that there exists a system-size parameter $K > 0$ such that the discrete jumps between states happen in units of $1/K$ and the total population size is controlled by $K$, with $K = \infty$ corresponding to an infinitely large population. In particular, we assume that the birth and death rates scale such that we can make the substitutions
\begin{align*}
\mathbf{x} &\coloneqq \frac{\mathbf{n}}{K}\\
b^{(K)}_i(\mathbf{x}) &\coloneqq \frac{1}{K}b_i(\mathbf{n})\\
d^{(K)}_i(\mathbf{x}) &\coloneqq \frac{1}{K}d_i(\mathbf{n})
\end{align*}
where $\mathbf{x} = \mathbf{n}/K$ measures population \emph{density} instead of population numbers. We now define new step operators $\Delta_{i}^{\pm}$ by their action on any real-valued function $f(\mathbf{x},t)$ as
\begin{equation}
\label{nD_step_operators_rescaled}
\Delta_{i}^{\pm}f([x_1,\ldots,x_m],t) = f([x_1,\ldots,x_i \pm \frac{1}{K}, \ldots x_m],t)
\end{equation}
In terms of these new variables, \eqref{nD_M_eqn} becomes
\begin{equation}
\label{nd_M_eqn_rescaled}
\frac{\partial P}{\partial t}(\mathbf{x},t) = K\sum\limits_{j=1}^{m}\left[(\Delta_j^{-}-1)b^{(K)}_j(\mathbf{x})P(\mathbf{x},t) + (\Delta_j^{+}-1)d^{(K)}_j(\mathbf{x})P(\mathbf{x},t)\right]
\end{equation}
If $K$ is large, we can once again Taylor expand the action of the step operators as
\begin{equation*}
f([x_1,\ldots,x_i \pm \frac{1}{K}, \ldots x_m],t) = f(\mathbf{x},t) \pm \frac{1}{K}\frac{\partial f}{\partial x_i}(\mathbf{x},t) + \frac{1}{2K^2}\frac{\partial^2f}{\partial x_i^2}(\mathbf{x},t) + \mathcal{O}(K^{-3})
\end{equation*}
which, after substituting into \eqref{nd_M_eqn_rescaled} and neglecting $ \mathcal{O}(K^{-3})$ terms, yields the equation
\begin{equation}
\label{nD_FPE}
\setlength{\fboxsep}{2\fboxsep}\boxed{\frac{\partial P}{\partial t}(\mathbf{x},t) = \sum\limits_{j=1}^{m}\left[-\frac{\partial}{\partial x_j}\{A_j^{-}(\mathbf{x})P(\mathbf{x},t)\} + \frac{1}{2K}\frac{\partial^2}{\partial x_j^2}\{A_j^{+}(\mathbf{x})P(\mathbf{x},t)\}\right]}
\end{equation}
where
\begin{equation*}
A_{i}^{\pm}(\mathbf{x}) = b^{(K)}_i(\mathbf{x})\pm d^{(K)}_i(\mathbf{x})
\end{equation*}
Equation \eqref{nD_FPE} is an $m$-dimensional Fokker-Planck equation, and corresponds to the $m$-dimensional It\^o process
\begin{equation}
\label{nD_Ito_SDE}
d\mathbf{X}_{t} = \mathbf{A^-}(\mathbf{X}_t)dt + \frac{1}{\sqrt{K}}\mathbf{D}(\mathbf{X}_t)d\mathbf{W}_t
\end{equation}
where $\mathbf{A^-}(\mathbf{X}_t)$ is an $m$-dimensional vector with $i$\textsuperscript{th} element $ = A^{-}_{i}(\mathbf{X}_t)$. $\mathbf{D}(\mathbf{X}_t)$ is an $m \times m$ matrix with $ij$th element $\left(\mathbf{D}(\mathbf{X}_t)\right)_{ij} = \delta_{ij}\left(A^{+}_{i}A^{+}_{j}\right)^{\frac{1}{4}}$, where $\delta_{ij}$ is the Kronecker delta symbol, defined by
\begin{equation*}
\delta_{ij} = 
\begin{cases}
1 & i=j\\
0 & i\neq j
\end{cases}
\end{equation*}
Finally, $\mathbf{W}_t$ is the $m$-dimensional Wiener process and can be thought of as a vector of independent one-dimensional Wiener processes (which have been defined in \ref{intro_SDE}). This is the `mesoscopic' description of our process.

@@

\section{Statistical measures for population-level quantities}\label{sec_stat_measures}
Given a state $\mathbf{x}(t)$ that describes our population at time $t$, we can compute the total (scaled) population size and the frequency of each type in the population as:
\begin{equation}
\label{nD_tot_pop_and_prop_inds_defn}
\begin{aligned}
N_{K}(t) &\coloneqq \sum\limits_{i=1}^{m}x_i(t) = \frac{1}{K}\sum\limits_{i=1}^{m}v_i(t)\\
p_i(t) &\coloneqq \frac{x_i(t)}{N_{K}(t)}
\end{aligned}
\end{equation}
These definitions are important because evolution is often defined via changes in type frequency (and \emph{not} necessarily changes in type density).

Even though the causes of evolution are generally at the individual level, we are very often interested in quantities \emph{described} at a population level, such as the mean fitness or mean phenotype in the population~\citep{bourrat_evolution_2019}. Furthermore, the relevant quantities at the individual level, such as individual fitness or phenotype, are typically equal for all individuals of the same type (in some sense this is our basis for defining different types in the first place). I use the term `type-level quantities' henceforth to refer to such quantities that are equal for all individuals that are of the same type.

Let $f$ be any type-level quantity with (possibly time-dependent) value $f_i$ for the $i\textsuperscript{th}$ type. For example, if each type is a phenotype for a trait such as height, which can be assigned a numerical value, then setting $f_i = \textit{value of $i\textsuperscript{th}$ phenotype}$ gives us the mean trait value in the population. To facilitate the description of such quantities, I define the statistical mean value of any quantity $f$ in the population as
\begin{equation}
\label{nD_mean}
\overline{f}(t) \coloneqq \sum\limits_{i=1}^{m}f_ip_{i}
\end{equation}
the statistical covariance of two such quantities $f$ and $g$ as
\begin{equation}
\label{nD_cov}
\textrm{Cov}(f,g) \coloneqq \overline{fg} - \overline{f}\overline{g}
\end{equation}
and the statistical variance of a quantity $f$ as $\sigma^2_f \coloneqq \textrm{Cov}(f,f)$.

@@

The first term of \eqref{nD_replicator_mutator} describes changes due to faithful (non-mutational) replication, and the second describes changes due to mutation. For this reason, equation \eqref{nD_replicator_mutator} is called the \emph{replicator-mutator equation} in the evolutionary game theory literature, where the individual `types' are interpreted to be pure strategies~\citep{hofbauer_evolutionary_1998,page_unifying_2002,cressman_replicator_2014}. If in addition, each $w_i(\mathbf{x})$ is linear in $\mathbf{x}$, meaning we can write $w_i(\mathbf{x}) = \sum_{j}a_{ij}x_j$ for some set of constants $a_{ij}$, then we get the replicator-mutator equation for matrix games, and the constants $a_{ij}$ form the `payoff matrix'. As is well-known, the replicator equation (without mutation) for matrix games with $m$ pure strategies is equivalent to the generalized Lotka-Volterra equations for a community with $m-1$ species~\citep{hofbauer_evolutionary_1998}, providing the connection to community ecology.  Equation \eqref{nD_replicator_mutator} is also equivalent to Eigen's \emph{quasispecies equation} from molecular evolution if each `type' is interpreted as a genetic sequence and each $w_i(\mathbf{x})$ is a constant function\footnote{Mutational effects are often additionally assumed to act through direct `transmission probabilities' of mutating from one type to another. This means that we can write $Q_i(\mathbf{p}) = \sum_j Q_{ij}p_j$, where $Q_{ii} = 0$, and for each $j \neq i$, $Q_{ij} \geq0$ is a constant describing the probability of a $j \to i$ mutation (conditioned on the occurrence of a mutation). Substituting this into \eqref{nD_replicator_mutator} yields an equation in terms of `$Q$-matrices' or `mutation matrices' that may be more familiar to some.}~\citep{page_unifying_2002}. We can now calculate how the mean of any `type level' quantity $f$, defined as $f_i$ for the $i$\textsuperscript{th} type, changes in the population (For example, if each type is a phenotype for a trait such as height, which can be assigned a numerical value, then setting $f_i = \textit{value of $i\textsuperscript{th}$ phenotype}$ gives us the mean trait value in the population). The product rule of calculus tells us that we have the relation

@@

A^-_{j}(\boldsymbol{\alpha} + \frac{\mathbf{y}}{\sqrt{K}}) = \sum\limits_{n=1}^{\infty}(A^{-}_{j})_{n}(s)\left(\frac{\mathbf{y}}{\sqrt{K}}\right)^n

>>>>

\section{Description of the process and the Master Equation}
Given a population that can contain up to $m$ different (fixed) kinds of organisms, it can be entirely characterized by specifying the number of organisms of each type (Figure \ref{fig_nD_pop_description}A). Thus, the state of the population at a given time $t$ is an $m$-dimensional \emph{vector} of the form $\mathbf{n}(t) = [n_1(t),n_2(t),\ldots,n_m(t)]^{\mathrm{T}}$, where $n_i(t)$ is the number of individuals of type $i$ in the population at time $t$.

@@

As before, we can describe the rate of change of $P(\mathbf{n},t)$, the probability of finding the population in a state $\mathbf{n}$ at time $t$, by measuring the inflow and outflow rates. Since the population changes in units of exactly one individual (by definition of a birth-death process; see section \ref{sec_math_background}), we know that these inflow and outflow rates must only involve populations that are a single individual away from our focal population. In other words, for a population $\mathbf{n} = [n_1,\ldots,n_{m}]^{\mathrm{T}}$, the `inflow' is from all populations of the form $[n_1,\ldots,n_{i}-1,\dots,n_{m}]^{\mathrm{T}}$ through a birth of a type $i$ individual, and from all populations of the form $[n_1,\ldots,n_{i}+1,\dots,n_{m}]^{\mathrm{T}}$ through the death of a type $i$ individual. Thus, we have the inflow rate
\begin{equation}
\label{nD_rate_in}
\begin{split}
R_{\textrm{in}}(\mathbf{n},t) &= \sum\limits_{j=1}^{m}b_{j}([n_1,\ldots,n_{j}-1,\ldots,n_m]^{\mathrm{T}})P([n_1,\ldots,n_{j}-1,\ldots,n_m]^{\mathrm{T}},t) \\
& +\sum\limits_{j=1}^{m}d_{j}([n_1,\ldots,n_{j}+1,\ldots,n_m]^{\mathrm{T}})P([n_1,\ldots,n_{j}+1,\ldots,n_m]^{\mathrm{T}},t)
\end{split}
\end{equation}
Outflow is through births and deaths of individuals in the population $\mathbf{n}$ itself, and thus we have:
\begin{equation}
\label{nD_rate_out}
R_{\textrm{out}}(\mathbf{n},t) = \sum\limits_{j=1}^{m}b_{j}(\mathbf{n})P(\mathbf{n},t) + \sum\limits_{j=1}^{m}d_{j}(\mathbf{n})P(\mathbf{n},t)
\end{equation}
We will now define step operators, both for notational ease and in anticipation of the system size expansion. For each $i \in \{1,\ldots,m\}$, let us define two step operators $\mathcal{E}_{i}^{\pm}$ by their action on any function $f([n_1,\ldots,n_m],t)$ as:
\begin{equation}
\label{nD_step_operators}
\mathcal{E}_{i}^{\pm}f([n_1,\ldots,n_i,\ldots,n_m]^{\mathrm{T}},t) = f([n_1,\ldots,n_i \pm 1, \ldots n_m]^{\mathrm{T}},t)
\end{equation}
In other words, $\mathcal{E}_{i}^{\pm}$ just changes the population through the addition or removal of one type $i$ individual. We can now the rate of change of $P(\mathbf{n},t)$ as
\begin{equation}
\label{nD_intermediate_for_M_eqn}
\frac{\partial P}{\partial t}(\mathbf{n},t) = R_{\textrm{in}}(\mathbf{n},t) - R_{\textrm{out}}(\mathbf{n},t)
\end{equation}
Substituting \eqref{nD_rate_in}, \eqref{nD_rate_out}, and \eqref{nD_step_operators} into equation \eqref{nD_intermediate_for_M_eqn}, we obtain:
\begin{equation}
\label{nD_M_eqn}
\frac{\partial P}{\partial t}(\mathbf{n},t) = \sum\limits_{j=1}^{m}\left[(\mathcal{E}_j^{-}-1)b_j(\mathbf{n})P(\mathbf{n},t) + (\mathcal{E}_j^{+}-1)d_j(\mathbf{n})P(\mathbf{n},t)\right]
\end{equation}
This is the master equation of our $m$-dimensional process.

\section{The system-size expansion}
As explained in Section \ref{sec_intuition_sys_size}, we will now assume (on ecological grounds) that there exists a system-size parameter $K > 0$ such that the discrete jumps between states happen in units of $1/K$ and the total population size is controlled by $K$, with $K = \infty$ corresponding to an infinitely large population. In particular, we assume that the birth and death rates scale such that we can make the substitutions
\begin{align*}
\mathbf{x} &\coloneqq \frac{\mathbf{n}}{K}\\
b^{(K)}_i(\mathbf{x}) &\coloneqq \frac{1}{K}b_i(\mathbf{n})\\
d^{(K)}_i(\mathbf{x}) &\coloneqq \frac{1}{K}d_i(\mathbf{n})
\end{align*}
where $\mathbf{x} = \mathbf{n}/K$ measures population \emph{density} instead of population numbers. We now define new step operators $\Delta_{i}^{\pm}$ by their action on any real-valued function $f(\mathbf{x},t)$ as
\begin{equation}
\label{nD_step_operators_rescaled}
\Delta_{i}^{\pm}f([x_1,\ldots,x_m]^{\mathrm{T}},t) = f([x_1,\ldots,x_i \pm \frac{1}{K}, \ldots x_m]^{\mathrm{T}},t)
\end{equation}
In terms of these new variables, \eqref{nD_M_eqn} becomes
\begin{equation}
\label{nd_M_eqn_rescaled}
\frac{\partial P}{\partial t}(\mathbf{x},t) = K\sum\limits_{j=1}^{m}\left[(\Delta_j^{-}-1)b^{(K)}_j(\mathbf{x})P(\mathbf{x},t) + (\Delta_j^{+}-1)d^{(K)}_j(\mathbf{x})P(\mathbf{x},t)\right]
\end{equation}
If $K$ is large, we can once again Taylor expand the action of the step operators as
\begin{equation*}
f([x_1,\ldots,x_i \pm \frac{1}{K}, \ldots x_m]^{\mathrm{T}},t) = f(\mathbf{x},t) \pm \frac{1}{K}\frac{\partial f}{\partial x_i}(\mathbf{x},t) + \frac{1}{2K^2}\frac{\partial^2f}{\partial x_i^2}(\mathbf{x},t) + \mathcal{O}(K^{-3})
\end{equation*}
which, after substituting into \eqref{nd_M_eqn_rescaled} and neglecting $ \mathcal{O}(K^{-3})$ terms, yields the equation
\begin{equation}
\label{nD_FPE}
\setlength{\fboxsep}{2\fboxsep}\boxed{\frac{\partial P}{\partial t}(\mathbf{x},t) = \sum\limits_{j=1}^{m}\left[-\frac{\partial}{\partial x_j}\{A_j^{-}(\mathbf{x})P(\mathbf{x},t)\} + \frac{1}{2K}\frac{\partial^2}{\partial x_j^2}\{A_j^{+}(\mathbf{x})P(\mathbf{x},t)\}\right]}
\end{equation}
where
\begin{equation*}
A_{i}^{\pm}(\mathbf{x}) = b^{(K)}_i(\mathbf{x})\pm d^{(K)}_i(\mathbf{x})
\end{equation*}
Equation \eqref{nD_FPE} is an $m$-dimensional Fokker-Planck equation, and corresponds to the $m$-dimensional It\^o process
\begin{equation}
\label{nD_Ito_SDE}
d\mathbf{X}_{t} = \mathbf{A^-}(\mathbf{X}_t)dt + \frac{1}{\sqrt{K}}\mathbf{D}(\mathbf{X}_t)d\mathbf{W}_t
\end{equation}
where $\mathbf{A^-}(\mathbf{X}_t)$ is an $m$-dimensional vector with $i$\textsuperscript{th} element $ = A^{-}_{i}(\mathbf{X}_t)$. $\mathbf{D}(\mathbf{X}_t)$ is an $m \times m$ matrix with $ij$th element $\left(\mathbf{D}(\mathbf{X}_t)\right)_{ij} = \delta_{ij}\left(A^{+}_{i}A^{+}_{j}\right)^{\frac{1}{4}}$, where $\delta_{ij}$ is the Kronecker delta symbol, defined by
\begin{equation*}
\delta_{ij} = 
\begin{cases}
1 & i=j\\
0 & i\neq j
\end{cases}
\end{equation*}
Finally, $\mathbf{W}_t$ is the $m$-dimensional Wiener process and can be thought of as a vector of independent one-dimensional Wiener processes (which have been defined in \ref{intro_SDE}). This is the `mesoscopic' description of our process.

@@

\section{Statistical measures for population-level quantities}\label{sec_stat_measures}

Though the causes of evolution are generally through ecological phenomena that occur at the \emph{individual} level like we have been working with, we are very often interested in tracking the effects of evolution on quantities described at a \emph{population} level, such as the mean fitness or mean phenotype in the population~\citep{bourrat_evolution_2019}. Furthermore, the relevant quantities at the individual level, such as individual fitness or phenotype, are typically equal for all individuals of the same type (in some sense this is our basis for defining different types in the first place). I use the term `type-level quantities' henceforth to refer to such quantities that are equal for all individuals that are of the same type. To facilitate the description of such quantities, given any state $\mathbf{x}(t)$ that describes our population at time $t$, let us first define the total (scaled) population size and the frequency of each type in the population as:
\begin{equation}
	\label{nD_tot_pop_and_prop_inds_defn}
	\begin{aligned}
		N_{K}(t) &\coloneqq \sum\limits_{i=1}^{m}x_i(t) = \frac{1}{K}\sum\limits_{i=1}^{m}n_i(t)\\
		p_i(t) &\coloneqq \frac{x_i(t)}{N_{K}(t)} = \frac{n_i(t)}{\sum\limits_{j=1}^{m}n_i(t)}
	\end{aligned}
\end{equation}
Now, let $f$ be any type-level quantity with (possibly time-dependent) value $f_i$ for the $i\textsuperscript{th}$ type. For example, if each type is a phenotype for a trait such as height, which can be assigned a numerical value, then setting $f_i = \textit{value of $i\textsuperscript{th}$ phenotype}$ gives us the mean trait value in the population. We can compute the statistical mean value of any such quantity in the population as
\begin{equation}
\label{nD_mean}
\overline{f}(t) \coloneqq \sum\limits_{i=1}^{m}f_ip_{i}
\end{equation}
the statistical covariance of two such quantities $f$ and $g$ as
\begin{equation}
\label{nD_cov}
\textrm{Cov}(f,g) \coloneqq \overline{fg} - \overline{f}\overline{g}
\end{equation}
and the statistical variance of a quantity $f$ as $\sigma^2_f \coloneqq \textrm{Cov}(f,f)$.

@@

The first term of \eqref{nD_replicator_mutator} describes changes due to faithful (non-mutational) replication, and the second describes changes due to mutation. For this reason, equation \eqref{nD_replicator_mutator} is called the \emph{replicator-mutator equation} in the evolutionary game theory literature, where the individual `types' are interpreted to be pure strategies~\citep{hofbauer_evolutionary_1998,page_unifying_2002,cressman_replicator_2014}. If in addition, each $w_i(\mathbf{x})$ is linear in $\mathbf{x}$, meaning we can write $w_i(\mathbf{x}) = \sum_{j}a_{ij}x_j$ for some set of constants $a_{ij}$, then we recover the more well-known replicator-mutator equation for matrix games in which the constants $a_{ij}$ form the `payoff matrix' (See the example presented in Appendix~\ref{App_examples}). As is well-known, the replicator equation (without mutation) for matrix games with $m$ pure strategies is equivalent to the generalized Lotka-Volterra equations for a community with $m-1$ species~\citep{hofbauer_evolutionary_1998}, providing the connection to community ecology.  Equation \eqref{nD_replicator_mutator} is also equivalent to Eigen's \emph{quasispecies equation} from molecular evolution if each `type' is interpreted as a genetic sequence and each $w_i(\mathbf{x})$ is a constant function\footnote{Mutational effects are often additionally assumed to act through direct `transmission probabilities' of mutating from one type to another. This means that we can write $Q_i(\mathbf{p}) = \sum_j Q_{ij}p_j$, where $Q_{ii} = 0$, and for each $j \neq i$, $Q_{ij} \geq0$ is a constant describing the probability of a $j \to i$ mutation (conditioned on the occurrence of a mutation). Substituting this into \eqref{nD_replicator_mutator} yields an equation in terms of `$Q$-matrices' or `mutation matrices' that may be more familiar to some.}~\citep{page_unifying_2002}. We can now calculate how the mean of any `type level' quantity $f$, defined as $f_i$ for the $i$\textsuperscript{th} type, changes in the population (For example, if each type is a phenotype for a trait such as height, which can be assigned a numerical value, then setting $f_i = \textit{value of $i\textsuperscript{th}$ phenotype}$ gives us the mean trait value in the population). The product rule of calculus tells us that we have the relation

@@

A^-_{j}(\boldsymbol{\alpha} + \frac{\mathbf{y}}{\sqrt{K}}) = \sum\limits_{n=0}^{\infty}(A^{-}_{j})_{n}(s)\left(\frac{\mathbf{y}}{\sqrt{K}}\right)^n
